Characterizing subarctic biomes for land surface modeling of pollution and climate risk

### 2021-03-03 VegOzone.tex
Change references accordingly to renamed files

\subsubsection{Ozone}
\label{subsubsec:clim_ozone}
In the following, we will look at derived ozone climatologies from different perspectives.
The multi-annual means of daily mean ozone, for simplicity here without standard deviations, at observation sites in northern Fennoscandia (refer to Table~\ref{tab:ebas_obs} for details) are shown in Supplement Fig.~S1. All climatologies peak in spring, with peak values of about $48\,\unit{ppb}$ in late April. The annual average ozone concentration \chem{\left<[O_3]\right>} declines throughout the growing season until it reaches its minimum in late summer/fall ($24-30\,\unit{ppb}$). The turning points lie approximately in June and November. The ozone climatology lags about 3 months behind the temperature climatology. At Svanvik, \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than on sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN). The large spread in case of Svanvik and Janiskoski can be explained due to the lower overall statistics.

@@

From the northern Fennoscandic climatology, one can deduce several features. The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} peaks about $20\,\unit{days}$ earlier than the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, and \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}}. With respect to the periodicity of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, it takes about $7.5\,\unit{months}$ from minimum concentration to maximum concentration, while from maximum to minimum it only takes about $4.5\,\unit{months}$. The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} is at $52\,\unit{ppb}$ in late April and the minimum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} occurs at $15\,\unit{ppb}$ in the beginning of September. The steep rise of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} in spring is in coincidence with increasing temperatures as well as increase in global irradiance. The decrease in \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} begins about one month ahead of the estimated onset of the greening season. Based on data from SeNorge.no and the agricultural rule of thumb that the daily average temperature has to be above $5\,\unit{^\circ C}$ for $5\,\unit{days}$ consecutively, we find $\left<G_\mathrm{begin}\right> = (150 \pm 14)\,\unit{days of the year}$ at Svanvik (Appendix Fig~\ref{fig:greening_season_change_Svanvik}), which translates into the end of May. This finding is in line with more extensive studies \citep[][e.g.,]{IJB:Karlsen2007,AFM:Linderholm}.

@@

\begin{figure}[t]
  \includegraphics[width=12cm]{fig08-11}
  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as percentages of days deviating from climatological mean by means of standart deviation for each month. Negative deviations from climatology are shown as negative percentage. The annual positive/negative deviations are indicated in the respective corners (right upper/right lower). The hatched area between the dashed lines indicates statistical significance under the assumption of a normal distribution ($15.9\,\unit{\%}$, $30.8\,\unit{\%}$ for precip.). (a) Temperature, days warmer/colder than $\pm 1\,\sigma$; (b) precipitation, days wetter/drier than $\pm \frac{1}{2}\,\sigma$; (c) global irradiance, days lighter/darker than $\pm 1\,\sigma$; (d) ozone, days higher/lower than $\pm 1\,\sigma$. The July 2018 ozone data corrected for data acquisition problems is indicated with the star (refer Section~\ref{subsec:ozone_reco} for details). The growing season of 2018 was significantly warmer, brighter, and slightly drier than average, especially in May and July. Ozone was significantly elevated in May. The growing season of 2019 was fairly average but significantly warmer than average in April and colder in July/August. No significant elevation in ozone was found.}
  \label{fig:anomalies_svanvik}
\end{figure}

@@

In Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}, the probability density functions of daily mean ozone concentration residuals for Esrange and Pallas are shown. Both stations display a clear deviation from the climatological mean, however, not significant. On average, \chem{[O_3]} had been elevated by $(1.9\pm 5.7)\,\unit{ppb}$ and $(1.9\pm 5.4)\,\unit{ppb}$, respectively. Both distributions are slightly skewed towards positive deviations with $\frac{1}{4}$ of days (3. quantile, q3) falling above $5.8\,\unit{ppb}$ and $5.6\,\unit{ppb}$, respectively.

Appendix Figure~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik} displays the probability density functions of daily mean ozone concentration residuals for Svanvik based on data taken in 2018/19 with respect to the climatology derived for northern Fennoscandia (denoted NF in Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}a) and the climatology derived for Svanvik (Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b).
As previously mentioned, \chem{\left<[O_3]\right>} at Svanvik are $(6.6\pm 7.1)\,\unit{ppb}$ lower than the northern Fennoscandic climatology. For 2018 and 2019, we find $\Delta\chem{\left<[O_3]\right>} = (-3.8\pm 5.6)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (-5.2\pm 5.0)\,\unit{ppb}$, respectively. While the probability density of the residuals between the two climatologies is skewed towards negative values, the skew of 2018/19 is positive.
With respect to the climatology for Svanvik (Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b), the probability density functions average for 2018/19 lies at $\Delta\chem{\left<[O_3]\right>} = (2.6\pm 5.8)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (1.2\pm 5.0)\,\unit{ppb}$, respectively.
Neither of the \chem{\Delta\left<[O_3]\right>} is significant by statistical means, but daily average ozone concentrations are slightly elevated in 2018 compared to 2019.

@@

To further investigate the ozone residuals, we conduct a Student's~t-test ($t = \frac{\chem{\Delta[O_3]}}{\sigma}$), assuming the same sample uncertainty $\sigma$ in both, climatology and the year in question. Results for Esrange and Pallas are shown in Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test}. This t-test clearly shows, that at both sites the residuals are significant on the $1\sigma$ level for at least $50\,\unit{\%}$. In Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}, the test statistics for Svanvik are displayed. From Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}a), we can deduce, that the residuals between the Svanvik climatology and the northern Fennoscandic climatology are significant on the $2\sigma$ level for at least half of the data, while the residuals are less certain in 2018 and 2019. In Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}b) the same is shown for the years 2018/19 with respect to the climatology from Svanvik. We find that less than $50\,\unit{\%}$ of ozone concentration residuals are significant on the $1\sigma$ level.

@@

Because the available reanalysis products for surface ozone have their limitations, we deem it necessary to reconstruct the missing data in 2018 for the purpose of $\mathrm{DO_3SE}$ modeling by applying a Reynolds's decomposition of the form:
\begin{equation}
  \chem{[O_3]} = \chem{\left<[O_3]\right>} + \chem{\Delta[O_3]},
  \label{eq:reynolds_decomp}
\end{equation}
with climatology \chem{\left<[O_3]\right>} and anomalies \chem{\Delta[O_3]}.
The exact procedure will be described in the following.
We derive hourly climatologies for individual stations and combine Esrange/Pallas to derive a climatology for northern Fennoscandia for July (Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}a). We compute temporal correlations between Svanvik and the other stations in northern Fennoscandia to estimate a time lag (Appendix Fig.~\ref{fig:time_lag_correlation}. We find $3\,\unit{h}$ for Esrange and Pallas with respect to Svanvik. Due to the high correlation with Svanvik (Appendix~Fig.~\ref{fig:density_distribution}), we decide on Pallas as reference station for the reconstruction. We correct the Fennoscandic climatology for the found time lag \chem{\left<[O_3]\right>_{hourly,\,t-3}} by shifting the respective time series back in time by $3\,\unit{h}$ and use it to derive a projection for Svanvik:
\begin{equation}
  X = \frac{\chem{\left<[O_3]\right>_{hourly}^{Svanvik}}}{\chem{\left<[O_3]\right>_{hourly,\,t-3}}}.
  \label{eq:mapping}
\end{equation}
As an intermediate step, we apply Eq.~(\ref{eq:reynolds_decomp}) to derive hourly anomalies for each station $i$
\begin{equation}
  \chem{\Delta[O_3]_{hourly}}^i = \chem{[O_3]_{hourly}}^i-\chem{\left<[O_3]\right>_{hourly}},
\end{equation}
with $i \in \{\text{Esrange, Pallas, Svanvik\}}$.
In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}b), these hourly ozone concentration anomalies are shown. We also correct the ozone anomalies for Pallas for the time lag \chem{\Delta[O_3]_{hourly,\,t-3}^{Pallas}} and use the mapping function in Eq.~(\ref{eq:mapping}) to derive reconstructed anomalies for the missing values at Svanvik:
\begin{equation}
    \chem{\Delta [O_3]_{hourly}^{Svanvik,\,reco}} = \chem{\Delta[O_3]_{hourly,\,t-3}^{Pallas}} \cdot X
\end{equation}
Finally, we add the bias corrected Svanvik ozone climatology and derive a reconstructed ozone concentration time series \chem{[O_3]_{hourly}^{Svanvik,\,reco}}. In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}c), the result is shown together with the observed data and the Copernicus Regional Model Reanalysis at the nearest grid point for comparison.

@@

For the acclimatization of the temperature response functions, we assume two different types: \emph{Boreal} and \emph{cold}. In either cases, we further assume that the hypothetical species are more tolerant to cold conditions than the mapping manual default suggests and shift $f_\mathrm{T}$ towards colder temperatures. We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to the mapping manual default. In the same way, \emph{Boreal} refers to a hypothetical species which is very tolerant to cold but is sensitive to high temperatures, and most efficient at cool temperatures (close to the climatological mean temperature of 1992--2000). All three types for grassland are depicted together with the temperature PDF in Fig.~\ref{fig:f_temp_grassland}a).

@@

\begin{figure}[t]
  %\includegraphics[width=8.3cm]{javis_funcs_temp_hist_grassland}\\
  %\includegraphics[width=8.3cm]{javis_funcs_rad_hist_grassland}
  \includegraphics[width=8.3cm]{fig12}
\caption{Construction of bespoke response functions for grassland. (a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively. Original mapping manual parameterization is shown in comparison as solid line. Note that $Q_0$ has been truncated to $0.006$. PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
\label{fig:f_temp_grassland}
\end{figure}

@@

\begin{figure}[t]
  %\includegraphics[width=8.3cm]{javis_func_opt_mean}
  \includegraphics[width=8.3cm]{fig13}
  \caption{Proposed metric to test bespoke response functions.}
  \label{fig:javis_func_opt_mean}
\end{figure}

Start ($A_\text{start}$) and end ($A_\text{end}$) of growing season for each PFT are given in units of day of year (doy) and estimated as follows. Evergreen needle leaf trees are photosynthetic active already at low temperatures \citep{TB:Kolari2007}. This leads necessarily to an increase in observable net photosynthesis. Hence, we use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine onset and end of growing season. Net photosynthesis data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}). By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{doy}$ for 2018 and day 106 in 2019. $A_\text{end}$ amounts to day 261 and 274, respectively, and shall be used for all vegetation alike. The resulting growing season for Norway spruce in 2019 was one month longer then in 2018. For estimating $A_\text{start}$ of birch tree, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}). We find 129 and 130, respectively. From the latitude model typically used for localizing $\mathrm{DO_3SE}$, we get $A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$ indicating an about $2\,\unit{months}$ longer growing season. For perennial grassland, we assume a latency period of about $1\,\unit{month}$ after snow melt. Data for first snow-free day is taken from the weather station at Øvre Neiden in Sør-Varanger. In both years, the first snow-free day incidents with the estimated date of bud burst. As leafs of birches in northern Fennoscandia \emph{Betula pubescens} are smaller than southern specimen, we have measured their dimension for the purpose of this study and find $(3.0\pm 0.5)\,\unit{cm}$. \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.

@@

\begin{figure}[t]
  %\includegraphics[width=8.3cm]{do3se_results_pod_v3_Birch}\\
  %\includegraphics[width=8.3cm]{do3se_results_pod_v3_Spruce}\\
  %\includegraphics[width=8.3cm]{do3se_results_pod_v3_Grassland}
  \includegraphics[width=8.3cm]{fig14}
  \caption{Comprehensive comparison of $\mathrm{DO_3SE}$ modeling results with default parameterization from mapping manual (MM)and bespoke parameters for Svanhovd. $\mathrm{POD_1}$ is shown for the different temperature (MM, boreal, cold) and light threshold (MM, PPFD0.8, PPFD1.2) parameterizations at the end of the MM GS (lighter colors and dotted lines) and a derived bespoke GS. (a) Birch; (b) Norway spruce; (c) Perennial grassland.}
  \label{fig:pody_rel}
\end{figure}

>>>>

\subsubsection{Ozone}
\label{subsubsec:clim_ozone}
In the following, we will look at derived ozone climatologies from different perspectives.
The multi-annual means of daily mean ozone, for simplicity here without standard deviations, at observation sites in northern Fennoscandia (refer to Table~\ref{tab:ebas_obs} for details) are shown in Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}. All climatologies peak in spring, with peak values of about $48\,\unit{ppb}$ in late April. The annual average ozone concentration \chem{\left<[O_3]\right>} declines throughout the growing season until it reaches its minimum in late summer/fall ($24-30\,\unit{ppb}$). The turning points lie approximately in June and November. The ozone climatology lags about 3 months behind the temperature climatology. At Svanvik, \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than on sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN). The large spread in case of Svanvik and Janiskoski can be explained due to the lower overall statistics.

@@

From the northern Fennoscandic climatology, one can deduce several features. The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} peaks about $20\,\unit{days}$ earlier than the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, and \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}}. With respect to the periodicity of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, it takes about $7.5\,\unit{months}$ from minimum concentration to maximum concentration, while from maximum to minimum it only takes about $4.5\,\unit{months}$. The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} is at $52\,\unit{ppb}$ in late April and the minimum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} occurs at $15\,\unit{ppb}$ in the beginning of September. The steep rise of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} in spring is in coincidence with increasing temperatures as well as increase in global irradiance. The decrease in \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} begins about one month ahead of the estimated onset of the greening season. Based on data from SeNorge.no and the agricultural rule of thumb that the daily average temperature has to be above $5\,\unit{^\circ C}$ for $5\,\unit{days}$ consecutively, we find $\left<G_\mathrm{begin}\right> = (150 \pm 14)\,\unit{days of the year}$ at Svanvik (Supplement Fig~\ref{fig:greening_season_change_Svanvik}), which translates into the end of May. This finding is in line with more extensive studies \citep[][e.g.,]{IJB:Karlsen2007,AFM:Linderholm}.

@@

\begin{figure}[t]
  \includegraphics[width=12cm]{fig08}
  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as percentages of days deviating from climatological mean by means of standart deviation for each month. Negative deviations from climatology are shown as negative percentage. The annual positive/negative deviations are indicated in the respective corners (right upper/right lower). The hatched area between the dashed lines indicates statistical significance under the assumption of a normal distribution ($15.9\,\unit{\%}$, $30.8\,\unit{\%}$ for precip.). (a) Temperature, days warmer/colder than $\pm 1\,\sigma$; (b) precipitation, days wetter/drier than $\pm \frac{1}{2}\,\sigma$; (c) global irradiance, days lighter/darker than $\pm 1\,\sigma$; (d) ozone, days higher/lower than $\pm 1\,\sigma$. The July 2018 ozone data corrected for data acquisition problems is indicated with the star (refer Section~\ref{subsec:ozone_reco} for details). The growing season of 2018 was significantly warmer, brighter, and slightly drier than average, especially in May and July. Ozone was significantly elevated in May. The growing season of 2019 was fairly average but significantly warmer than average in April and colder in July/August. No significant elevation in ozone was found.}
  \label{fig:anomalies_svanvik}
\end{figure}

@@

In Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}, the probability density functions of daily mean ozone concentration residuals for Esrange and Pallas are shown. Both stations display a clear deviation from the climatological mean, however, not significant. On average, \chem{[O_3]} had been elevated by $(1.9\pm 5.7)\,\unit{ppb}$ and $(1.9\pm 5.4)\,\unit{ppb}$, respectively. Both distributions are slightly skewed towards positive deviations with $\frac{1}{4}$ of days (3. quantile, q3) falling above $5.8\,\unit{ppb}$ and $5.6\,\unit{ppb}$, respectively.

Supplement Figure~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik} displays the probability density functions of daily mean ozone concentration residuals for Svanvik based on data taken in 2018/19 with respect to the climatology derived for northern Fennoscandia (denoted NF in Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}a) and the climatology derived for Svanvik (Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b).
As previously mentioned, \chem{\left<[O_3]\right>} at Svanvik are $(6.6\pm 7.1)\,\unit{ppb}$ lower than the northern Fennoscandic climatology. For 2018 and 2019, we find $\Delta\chem{\left<[O_3]\right>} = (-3.8\pm 5.6)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (-5.2\pm 5.0)\,\unit{ppb}$, respectively. While the probability density of the residuals between the two climatologies is skewed towards negative values, the skew of 2018/19 is positive.
With respect to the climatology for Svanvik (Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b), the probability density functions average for 2018/19 lies at $\Delta\chem{\left<[O_3]\right>} = (2.6\pm 5.8)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (1.2\pm 5.0)\,\unit{ppb}$, respectively.
Neither of the \chem{\Delta\left<[O_3]\right>} is significant by statistical means, but daily average ozone concentrations are slightly elevated in 2018 compared to 2019.

@@

To further investigate the ozone residuals, we conduct a Student's~t-test ($t = \frac{\chem{\Delta[O_3]}}{\sigma}$), assuming the same sample uncertainty $\sigma$ in both, climatology and the year in question. Results for Esrange and Pallas are shown in Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test}. This t-test clearly shows, that at both sites the residuals are significant on the $1\sigma$ level for at least $50\,\unit{\%}$. In Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}, the test statistics for Svanvik are displayed. From Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}a), we can deduce, that the residuals between the Svanvik climatology and the northern Fennoscandic climatology are significant on the $2\sigma$ level for at least half of the data, while the residuals are less certain in 2018 and 2019. In Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}b) the same is shown for the years 2018/19 with respect to the climatology from Svanvik. We find that less than $50\,\unit{\%}$ of ozone concentration residuals are significant on the $1\sigma$ level.

@@

Because the available reanalysis products for surface ozone have their limitations, we deem it necessary to reconstruct the missing data in 2018 for the purpose of $\mathrm{DO_3SE}$ modeling by applying a Reynolds's decomposition of the form:
\begin{equation}
  \chem{[O_3]} = \chem{\left<[O_3]\right>} + \chem{\Delta[O_3]},
  \label{eq:reynolds_decomp}
\end{equation}
with climatology \chem{\left<[O_3]\right>} and anomalies \chem{\Delta[O_3]}.
The exact procedure will be described in the following.
We derive hourly climatologies for individual stations and combine Esrange/Pallas to derive a climatology for northern Fennoscandia for July (Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}a). We compute temporal correlations between Svanvik and the other stations in northern Fennoscandia to estimate a time lag (Supplement Fig.~\ref{fig:time_lag_correlation}. We find $3\,\unit{h}$ for Esrange and Pallas with respect to Svanvik. Due to the high correlation with Svanvik (Appendix~Fig.~\ref{fig:density_distribution}), we decide on Pallas as reference station for the reconstruction. We correct the Fennoscandic climatology for the found time lag \chem{\left<[O_3]\right>_{hourly,\,t-3}} by shifting the respective time series back in time by $3\,\unit{h}$ and use it to derive a projection for Svanvik:
\begin{equation}
  X = \frac{\chem{\left<[O_3]\right>_{hourly}^{Svanvik}}}{\chem{\left<[O_3]\right>_{hourly,\,t-3}}}.
  \label{eq:mapping}
\end{equation}
As an intermediate step, we apply Eq.~(\ref{eq:reynolds_decomp}) to derive hourly anomalies for each station $i$
\begin{equation}
  \chem{\Delta[O_3]_{hourly}}^i = \chem{[O_3]_{hourly}}^i-\chem{\left<[O_3]\right>_{hourly}},
\end{equation}
with $i \in \{\text{Esrange, Pallas, Svanvik\}}$.
In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}b), these hourly ozone concentration anomalies are shown. We also correct the ozone anomalies for Pallas for the time lag \chem{\Delta[O_3]_{hourly,\,t-3}^{Pallas}} and use the mapping function in Eq.~(\ref{eq:mapping}) to derive reconstructed anomalies for the missing values at Svanvik:
\begin{equation}
    \chem{\Delta [O_3]_{hourly}^{Svanvik,\,reco}} = \chem{\Delta[O_3]_{hourly,\,t-3}^{Pallas}} \cdot X
\end{equation}
Finally, we add the bias corrected Svanvik ozone climatology and derive a reconstructed ozone concentration time series \chem{[O_3]_{hourly}^{Svanvik,\,reco}}. In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}c), the result is shown together with the observed data and the Copernicus Regional Model Reanalysis at the nearest grid point for comparison.

@@

For the acclimatization of the temperature response functions, we assume two different types: \emph{Boreal} and \emph{cold}. In either cases, we further assume that the hypothetical species are more tolerant to cold conditions than the mapping manual default suggests and shift $f_\mathrm{T}$ towards colder temperatures. We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to the mapping manual default. In the same way, \emph{boreal} refers to a hypothetical species which is very tolerant to cold but is sensitive to high temperatures, and most efficient at cool temperatures ($T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000). All three types for grassland are depicted together with the temperature PDF in Fig.~\ref{fig:f_temp_grassland}a).

@@

\begin{figure}[t]
  %\includegraphics[width=8.3cm]{javis_funcs_temp_hist_grassland}\\
  %\includegraphics[width=8.3cm]{javis_funcs_rad_hist_grassland}
  \includegraphics[width=8.3cm]{fig09}
\caption{Construction of bespoke response functions for grassland. (a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively. Original mapping manual parameterization is shown in comparison as solid line. Note that $Q_0$ has been truncated to $0.006$. PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
\label{fig:f_temp_grassland}
\end{figure}

@@

\begin{figure}[t]
  %\includegraphics[width=8.3cm]{javis_func_opt_mean}
  \includegraphics[width=8.3cm]{fig10}
  \caption{Proposed metric to test bespoke response functions.}
  \label{fig:javis_func_opt_mean}
\end{figure}

Start ($A_\text{start}$) and end ($A_\text{end}$) of growing season for each PFT are given in units of day of year (doy) and estimated as follows. Evergreen needle leaf trees show photosynthetic activity already at low temperatures \citep{TP:Wallin2013, TB:Kolari2007}. We use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine onset and end of growing season. Net photosynthesis data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}). By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{doy}$ for 2018 and day 106 in 2019. $A_\text{end}$ amounts to day 261 and 274, respectively, and shall be used for all vegetation alike. The resulting growing season for Norway spruce in 2019 was one month longer then in 2018. For estimating $A_\text{start}$ of birch tree, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}). We find 129 and 130, respectively. From the latitude model typically used for localizing $\mathrm{DO_3SE}$, we get $A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$ indicating an about $2\,\unit{months}$ longer growing season. For perennial grassland, we assume a latency period of about $1\,\unit{month}$ after snow melt. Data for first snow-free day is taken from the weather station at Øvre Neiden in Sør-Varanger. In both years, the first snow-free day incidents with the estimated date of bud burst. As leafs of birches in northern Fennoscandia \emph{Betula pubescens} are smaller than southern specimen, we have measured their dimension for the purpose of this study and find $(3.0\pm 0.5)\,\unit{cm}$. \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.

@@

\begin{figure}[t]
  %\includegraphics[width=8.3cm]{do3se_results_pod_v3_Birch}\\
  %\includegraphics[width=8.3cm]{do3se_results_pod_v3_Spruce}\\
  %\includegraphics[width=8.3cm]{do3se_results_pod_v3_Grassland}
  \includegraphics[width=8.3cm]{fig11}
  \caption{Comprehensive comparison of $\mathrm{DO_3SE}$ modeling results with default parameterization from mapping manual (MM)and bespoke parameters for Svanhovd. $\mathrm{POD_1}$ is shown for the different temperature response functions (MM, boreal, cold) and light thresholds (MM, PPFD0.8, PPFD1.2) at the end of a MM GS (lighter hue and dotted lines) and a derived bespoke GS. (a) Birch; (b) Norway spruce; (c) Perennial grassland.}
  \label{fig:pody_rel}
\end{figure}
